# Reproducible Research: Peer Assessment 1

# 
## Loading and preprocessing the data
# 

```{r "LoadData"}
data = read.csv("activity.csv")
data_by_date <- aggregate(data$steps,by=list(data$date),FUN=mean)
colnames(data_by_date) <- c("Date","Mean_Steps")
steps_by_date <- aggregate(data$steps,by=list(data$date),FUN=sum)
colnames(steps_by_date) <- c("Date","TotalSteps")

```

# 
## What is mean total number of steps taken per day?
# 

```{r "TotalNumberOfStepsPerDay", fig.width=7, fig.height=6}
library(ggplot2)
ggplot(steps_by_date, aes(x=TotalSteps)) + geom_histogram(binwidth = 1000) + ggtitle("Histogram - Total number of steps taken each day")
summary(steps_by_date)
meanTotalSteps = mean(steps_by_date$TotalSteps,na.rm = TRUE)
medianTotalSteps = median(steps_by_date$TotalSteps,na.rm = TRUE)
```
### ------------------
### MEAN   - Number of total steps per day  = `r meanTotalSteps`
### MEDIAN - Number of total steps per day  = `r medianTotalSteps`
### ------------------
# 
## What is the average daily activity pattern?
# 

```{r "DailyActivity", fig.width=7, fig.height=6}
data_by_interval <- aggregate(data$steps,by=list(data$interval),FUN=mean, na.rm = TRUE)
colnames(data_by_interval) <- c("Interval","Mean_Steps")
plot(data_by_interval$Interval, data_by_interval$Mean_Steps, type = "l", xlab = "Interval", ylab = "Mean Steps", main = "Daily Activity Pattern")
mm = max(data_by_interval$Mean_Steps)
sm = subset(data_by_interval,data_by_interval$Mean_Steps == max(data_by_interval$Mean_Steps))

```
### ------------------
### The `r sm$Interval`th 5-minute interval has maximum average steps `r mm` 
### ------------------
# 
## Imputing missing values
### Problem3 - step1 - Total number of rows with NAs
# 

```{r "MissingValues", fig.width=7, fig.height=6}
na_data <- subset(data, is.na(data$steps))
NumOfMissingValues = length(na_data$steps)
```
### ------------------
### The total number rows with missing values = `r NumOfMissingValues`
### ------------------
# 
### Problem3 - step2 - Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc
# 
### ------------------
### The strategy is to fill the missing values with mean of that 5 minute interval,
### ------------------
# 
### Problem3 - step3 - Create a new dataset that is equal to the original dataset but with the missing data filled in
# 

```{r "ReplaceMissingValues", fig.width=7, fig.height=6}
for (k in data_by_interval$Interval) {
subdata = subset(data,data$interval==k)
subdata$steps[is.na(subdata$steps)] <- data_by_interval$Mean_Steps[data_by_interval$Interval==k]
data$steps[data$interval==k] <- subdata$steps
}
```
# 
### Problem3 - step4 - Make a histogram of the total number of steps each day and calculate and report the mean and median total number of steps taken per day
# 

```{r "HistogramAfterReplacing", fig.width=7, fig.height=6}

steps_by_date <- aggregate(data$steps,by=list(data$date),FUN=sum)
colnames(steps_by_date) <- c("Date","TotalSteps")

library(ggplot2)
ggplot(steps_by_date, aes(x=TotalSteps)) + geom_histogram(binwidth = 1000) + ggtitle("Histogram - Total number of steps taken each day \n with missing values replaced")
summary(steps_by_date)
mean_TSteps = mean(steps_by_date$TotalSteps)       # mean number of steps per day
median_TSteps = median(steps_by_date$TotalSteps)   # median number of steps per day
```
### ------------------
### MEAN   - Number of total steps per day  = `r mean_TSteps`
### MEDIAN - Number of total steps per day  = `r median_TSteps`
### The mean stayed the same as before and median just has gone up by 1
### ------------------
# 
## Are there differences in activity patterns between weekdays and weekends? 
# 

```{r "WeekdayVSWeekend", fig.width=7, fig.height=6}

wk=transform(data, weekend=as.POSIXlt(date, origin='%Y-%m-%d')$wday %in% c(0, 6))
wk1=transform(wk,day="WeekDay")
wk1$day = ifelse(wk1$weekend == "TRUE", "WeekEnd", "WeekDay")


wk10 = subset(wk1,wk1$day=="WeekDay")
wk11 = subset(wk1,wk1$day=="WeekEnd")
wkday_by_interval <- aggregate(wk10$steps,by=list(wk10$interval),FUN=mean, na.rm = TRUE)
colnames(wkday_by_interval) <- c("interval","MeanSteps")
wkend_by_interval <- aggregate(wk11$steps,by=list(wk11$interval),FUN=mean, na.rm = TRUE)
colnames(wkend_by_interval) <- c("interval","MeanSteps")

wkend1 = cbind(wkend_by_interval,day="weekend")
wkday1 = cbind(wkday_by_interval,day="weekday")
wkdayend1 = rbind(wkday1,wkend1)
library(lattice)
xyplot(MeanSteps~interval|day, data = wkdayend1, ylab = "Number of steps",
panel = "panel.lines", layout = c(1,2),
scales = list(y = list(relation = "same")))

```

